---
title: "RegressionModelsProject"
output: html_document
date: "2024-07-01"
---
## mtcars dataset
A data frame with 32 observations on 11 (numeric) variables
[, 1]	mpg	Miles/(US) gallon
[, 2]	cyl	Number of cylinders
[, 3]	disp	Displacement (cu.in.)
[, 4]	hp	Gross horsepower
[, 5]	drat	Rear axle ratio
[, 6]	wt	Weight (1000 lbs)
[, 7]	qsec	1/4 mile time
[, 8]	vs	Engine (0 = V-shaped, 1 = straight)
[, 9]	am	Transmission (0 = automatic, 1 = manual)
[,10]	gear	Number of forward gears
[,11]	carb	Number of carburetors

```{r}
 carsdf <- data.frame(mtcars)
head(carsdf)
y<-carsdf$mpg
x<-carsdf$am
idx<-x==0

plot(x, y,xlab="Transmission",ylab="MilesPerGallon" ,type = "n", frame = FALSE)
points(x[idx], y[idx], pch = 21, col = "black", bg = "lightblue", cex = 2)
points(x[!idx], y[!idx], pch = 21, col = "black", bg = "salmon", cex = 2)

```

## Linear Regression 
the linear regression shows that changing transmission type from automatic to manual increases 7.24 miles per gallon.
```{r}
lmmpgtr <- lm(y~factor(x))
summary(lmmpgtr)$coef

```
 ## Linear Regression with weight as possible confounder
 
```{r}
fit2<-lm(y~factor(x)+carsdf$wt)
summary(fit2)$coef

```
  ##  model comparison
  as pvalue for the model comparison is less than 0.05 so it concludes that the second model (which is including weight) is better fit for our data.
```{r}
fit2<-lm(y~factor(x)+carsdf$wt)
anova(lmmpgtr,fit2)

```
 